### 2021-08-15 doc/thesis/chapters/dsp.tex
adapt all the diagrams
minor modifications

The superheterodyne receiver is often referred to as \textit{superhet}, or \textit{super} in the literature.
The main characteristic of a superhet is, that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\

@@

The depicted block diagram represents a superhet receiver structure, for example an FM broadcast receiver, where an audio output is generated from the antenna input.

@@

In order to achieve direct down-conversion, the local oscillator of the mixer needs to be set exactly to the center frequency of the desired signal.
The main advantage of direct-conversion receivers is, that their output is an image-free signal.
This is due to the properties of the mixer.
The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer's output, and the flicker noise, also called \textit{1/f}-noise, which induces strong noise at low frequencies.
These effects are not explained into more detail here, since this is not the focus of this thesis.

@@

A block diagram, as well as the signal spectrum of the process is shown in Figure~\ref{fig:bd_down_conversion_spectrum}. In this example, the frequency of the LO is set to shift the signal from $f_m$ to baseband.

@@

The procedure that is used for down-conversion is an amplitude modulation (AM).
The equation for a generic AM signal is $ S_{AM}(t) = A_m m(t) \cdot A_c c(t) $, where $m(t)$ represents the message signal, and $c(t)$ is the high-frequency carrier.
Their respective amplitudes are $A_m$ and $A_c$.
This equation is depicted as a block diagram and signal spectrum in Figure~\ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
The signal $m(t)$ is a baseband signal.
It may be antenna signal that was filtered by a bandpass filter.
The received signal is a real signal, and thus, its negative frequency spectrum is an exact mirror of the positive frequencies.
This fact can be explained in two simple equations.\\

@@

Sample rate reduction is also called downsampling.
Downsampling is a way to reduce the amount of data that needs to be processed, without losing any of the signals' information.
In some applications, the digital signal is oversampled by a large factor.
However, if the frequency spectrum of the desired signal is a narrow band, a high oversampling rate is not necessary and the signal can be downsampled without losing any information content.
To guarantee this, the Nyquist sampling theorem must still be met after the downsampling process.

@@

As an example, assuming the signal of interest has a bandwidth of 20~kHz.
The Nyquist theorem thus requires a remaining sampling rate of at least twice the signals' frequency, so 40~kHz.
The signal is sampled with an oversampling factor of 10 -- meaning a sample rate of 200~kHz.
A downsampling factor of 4 is now applied, which results in a sample rate of 50~kHz.
The resulting signal is now downsampled and can be processed at this lower data rate, while none of the original data content is lost.\\

@@

More detailed information can be found in \cite[Ch.~6.9]{AlessioSilviaMaria2016DSPa}, \cite[Ch.~4.1]{GopiE.S2018MDSP} and \cite[Ch.~10.2.2]{ThyagarajanK.S2019ItDS}.

@@

\begin{equation}
  \begin{split}
    (L+R) + (L-R) = (2)L \\
    (L+R) - (L-R) = (2)R
    \label{equ_stereo_from_sum_diff}
  \end{split}
\end{equation}

where $(L+R)$ represents the mono part and $(L-R)$ the stereo difference.\\

>>>>

The superheterodyne receiver is often referred to as \textit{superhet}, or \textit{super} in the literature.
The main characteristic of a superhet is that it converts the radio-frequency (RF) antenna signal to an IF in a first stage, instead of going to baseband directly.\\

@@

The depicted block diagram represents a superhet receiver structure, for example as it can be used in an FM broadcast receiver, where an audio output is generated from the antenna input.

@@

In order to achieve direct down-conversion, the local oscillator of the mixer needs to be set exactly to the center frequency of the desired signal.
The main advantage of direct-conversion receivers is, that their output is an image-free signal.
This is due to the properties of the mixer.
The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer's output, and the flicker noise, also called \textit{1/f}-noise, which induces strong noise at low frequencies.
These effects are not explained into more detail here, since this is not the focus of this thesis.
Figure~\ref{fig:bd_direct_conversion} shows a block diagram of a direct-conversion receiver.

@@

A block diagram, as well as the signal spectrum of the process is shown in Figure~\ref{fig:bd_down_conversion_spectrum}. In this example, the frequency of the LO is set to shift the signal from $f_m$, which is drawn in solid lines, to baseband, which is drawn in dashed lines.

@@

The equation for a generic AM signal is $ S_{AM}(t) = A_m m(t) \cdot A_c c(t) $, where $m(t)$ represents the message signal, and $c(t)$ is the high-frequency carrier.
Their respective amplitudes are $A_m$ and $A_c$.
This equation is depicted as a block diagram and signal spectrum in Figure~\ref{fig:bd_down_conversion_spectrum}, in which $S_{AM}$ is denoted as the baseband signal $m_{BB}$, because it is already down-converted.
The original HF signal $m$, before the down-conversion, is a bandpass signal, such as an antenna signal that was filtered by a bandpass filter.
This received signal is a real signal, and thus, its negative frequency spectrum is an exact mirror of the positive frequencies.
This fact can be explained in two simple equations.\\

@@

Sample rate reduction is also called downsampling.
It is a way to reduce the amount of data that needs to be processed, without losing any of the signals' information.
In some applications, the digital signal is oversampled by a large factor.
However, if the frequency spectrum of the desired signal is a narrow band, a high oversampling rate is not necessary and the signal can be downsampled without losing any information content.
To guarantee this, the Nyquist sampling theorem must still be met after the downsampling process.\\

@@

As an example, assuming the signal of interest has a bandwidth of 20~kHz.
The Nyquist theorem thus requires a remaining sampling rate of at least twice the signal's frequency, so 40~kHz.
The signal is sampled with an oversampling factor of 10 -- meaning a sample rate of 200~kHz.
A downsampling factor of 4 is now applied, which results in a sample rate of 50~kHz.
The resulting signal is now downsampled and can be processed at this lower data rate, while none of the original data content is lost.\\

@@

More detailed information can be found in the literature \cite[Ch.~6.9]{AlessioSilviaMaria2016DSPa}, \cite[Ch.~4.1]{GopiE.S2018MDSP} and \cite[Ch.~10.2.2]{ThyagarajanK.S2019ItDS}.

@@

\begin{equation}
  \begin{split}
    (L+R) + (L-R) = (2)L \\
    (L+R) - (L-R) = (2)R,
    \label{equ_stereo_from_sum_diff}
  \end{split}
\end{equation}

\noindent
where $(L+R)$ represents the mono part and $(L-R)$ the stereo difference.\\
